---
title: "SF, LR AI DW | 24H WD vs Control 1 groups: Bayesian regression of smoothed stretch forces"
author: "Yaromir Kobikov <kobikov.yaromir@gmail.com>"
date: "Date: `r format(Sys.time(), '%d/%m/%Y')`"
output:
  word_document: default
  html_document:
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}

source("../../BayesianPValue.R")

```

 # Read SF data 

# Contra vs ipsi

> **Stretch‑force experiments with symmetric WD|Control1 groups**  
> Force–distance curves from both hind limbs were recorded simultaneously after bilateral peptide injection.  
> Mechanical work for each limb (WL, WR) was obtained by integrating a loess‑smoothed force curve (span = 0.4, family = "symmetric") over 0–10 mm stretch.  
> Asymmetry was quantified as the work difference ΔW = WL − WR and the log‑ratio AI<sub>L/R</sub> = log₂(WL / WR).  

```{r input}

my.seed <- 20231103 # reproducible Bayessian fit
my.span <- 0.4

rds_folder_name <- "../../rds/sf/sf_0-2_LR_AI/"
# rds_folder_name <- "../../rds/sf/sf_0-2_CI_AI/"

dir.create(file.path(rds_folder_name), showWarnings = FALSE)
load("../SF Sp 0-1 0.4-1 1-2 0-2 0.4-2 loess(symmetric span0.4)_rats.RData")
rm(sf) # conserve memory, we do not need the raw data


dta <- sf_LR_AI %>%
# dta <- sf_CI_AI %>%
  filter(dT %in% c("T_0_2")) %>%
  rename(Time = Period) %>%
  droplevels(.)

```

The median (bars) and 95% HPDCI (whiskers) for SF AI data points. __NOT__ the uncertainties estimated by Bayesian model.

```{r plot_bars_hpdci_lr}

myname <- "SF 0-2 span=0.4 median and 95% HPDCI all group"

# adapted from http://www.cookbook-r.com/Graphs/Plotting_means_and_error_bars_(ggplot2)/
plt[["all_group"]] <- dta %>% 
  group_by(Trt, Time) %>% 
  median_hdci(AI, .width = 0.95) %>%
  ungroup() %>%
  ggplot(aes(x=interaction(Time, sep=", ", drop=TRUE), y=AI, fill=Trt)) + 
    geom_bar(position=position_dodge(), stat="identity") +
    geom_errorbar(aes(ymin=.lower, ymax=.upper),
                  width=.2,   # Width of the error bars
                  position=position_dodge(.9)) +
    geom_hline(yintercept=0, linetype="dashed", color="black") +
    theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
    labs(title=myname, x="Medians and 95% HDCI", y="AI")

print(plt[["all_group"]])

```

# Analyse Groups, time

The group sizes are (distinct rats only, same rats with different stimulation locations are counted as duplicates)

```{r input2_pre, echo=FALSE}
dta_pre <- dta %>% 
  filter(Time == '24H') %>%
  filter(Trt %in% c("24H WD", "Control 1")) %>% 
  droplevels(.)

d.sum <- dta_pre %>% 
  arrange(Trt, Time) %>% 
  group_by(Trt, Time) %>% 
  summarise(rats = length(unique(RatID)) ) %>%
  pivot_wider(names_from=c(Trt), values_from=rats)

d.sum %>% flextable %>% fontsize(part = "header", size = 9) %>% autofit

```

## SF AI: Bayesian regression, student family

Prior predictive check


```{r sf_check_prior_pre}

std.prior <- brm(
  data = dta_pre,
  family = gaussian,
  AI ~ 0 + Trt + (1|RatID),
  prior = c(
    prior(exponential(1),      class = "sd"),
    prior(student_t(3, 0, 2.5), class = "sigma"),
    prior(normal(0, 5),        class = "b")
  ),
  
  # prior = c(prior(normal(0, 3), class = "sigma"),
  #           prior(normal(0, 3), class = "sd"),
  #           prior(normal(0, 10), class = "b")),
  seed = my.seed, 
  chains = my.cores, 
  cores = my.cores, 
  iter = 4e4,
  sample_prior = "only",
  file = paste0(rds_folder_name, "_check_prior.rds"),
  file_refit = "on_change"
)


cond_eff <- conditional_effects(std.prior,
                                  "Trt",
                                  method="fitted"
                                )

plt[["SF Trt prior check"]] <- plot(cond_eff,
                                       points = TRUE,
                                       point_args = list(width = 0.1, shape="o", col="red"),
                                       plot=FALSE)[[1]] +
                                  geom_hline(yintercept=0, linetype="dashed", color="black")

set.seed(2)
plt[["SF Trt prior pp_check"]] <- pp_check(std.prior, ndraws = 11)

print(cond_eff[[1]])
print(plt[["SF Trt prior check"]])
print(plt[["SF Trt prior pp_check"]])

```

Model fit and QC:

```{r sf_model_summary_pre, echo=FALSE, fig.width = 10}

std.fit <-  update(std.prior, 
                   sample_prior = "yes",
                   # control = list(adapt_delta = 0.99),
                   # control = list(max_treedepth = 15),
                   seed = my.seed, 
                   cores = my.cores,
                   file = paste0(rds_folder_name, "model_summary.rds"),
                   file_refit = "on_change")
summary(std.fit)

print(plot(conditional_effects(std.fit), ask=FALSE, plot=FALSE))

```

## SF AI: MCMC conversion diagnostics

Posterior predictive check

```{r sf_MCMC_ppcheck_pre}

set.seed(2) # reproducible pp_check

plt[["SF Trt pp_check"]] <- pp_check(std.fit, ndraws = 21)
print(plt[["SF Trt pp_check"]])

```

Autocorrelations

```{r sf_MCMC_ac_pre}

plt[["SF Trt ac"]] <- stan_ac(std.fit$fit)
print(plt[["SF Trt ac"]])

```

## SF AI: Estimated model marginals

Estimated model medians +- 95% HPDCI (HPDCI = highest posterior density continuous interval) as a table:

```{r sf_emm_pre}

emm <- emmeans(std.fit, ~ Trt, regrid = "response", nesting = NULL)
emm_show(emm)

```

Estimated model means as a median +- 95% HPDCI (HPDCI = highest posterior density continuous interval):

```{r sf_emm2_pre, echo=FALSE, fig.width = 8}

plt[["SF Trt emm"]] <- emm %>% emm_inorder %>%
  ggplot(aes(x = interaction(Trt, sep=", ", drop=TRUE), y = .value, fill = Trt)) %>%
  my.stat_eye() +
  geom_hline(yintercept=0, linetype="dashed", color="black") +
  labs(title=myname, x="Estimated Marginal Medians and 95% HPDCI", y="AI")

```

```{r points_sf_emm2_pre, echo=FALSE, fig.width=8}

plt[["SF Trt emm with Data Points"]] <- 
  plt[["SF Trt emm"]] +
    geom_point(
      pch = 4,
      position = position_jitterdodge(jitter.width = 0.9, jitter.height = 0, dodge.width = 0.9, seed = 2),
      data = {
        dta_pre %>%
          group_by(RatID, Trt) %>%
          summarise(.value = mean(AI), .groups = "drop")
      },
      aes(y=.value))

print(plt[["SF Trt emm with Data Points"]])

```

The bar and whiskers plot:

```{r sf_bar_whiskers_pre, echo=FALSE}

plt[["SF Trt bar whiskers"]] <- emm %>% gather_emmeans_draws %>% 
  median_hdci(.width = 0.95) %>% 
  ungroup() %>%
  ggplot(aes(x = interaction(Trt, sep=", ", drop=TRUE), y = .value, fill = Trt)) + 
    geom_bar(position=position_dodge(), aes(y=.value),stat="identity") +
    geom_errorbar(aes(ymin=.lower, ymax=.upper),
                  width=.2,   # Width of the error bars
                  position=position_dodge(.9)) +
    geom_hline(yintercept=0, linetype="dashed", color="black") +
    theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
    labs(title=myname, x="Medians and 95% HDCI", y="AI")

print(plt[["SF Trt bar whiskers"]])

```

The ccdinterval plot with same data as above bars and whiskers:

```{r sf_ccdfinterval_pre, echo=FALSE}

plt[["SF Trt ccdfinterval"]] <- emm %>% gather_emmeans_draws %>%
  ggplot(aes(x = interaction(Trt, sep=", ", drop=TRUE), y = .value, fill = Trt)) +
  stat_ccdfinterval(point_interval = median_hdci, .width = 0.95,
             shape = 16, point_color = "black", interval_color = "black", 
             position="dodge", slab_color = NA) +
    scale_fill_brewer(palette = "Set2") + 
    scale_color_brewer(palette = "Dark2") +
    theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  # coord_cartesian(ylim = c(0, 15.5), expand=FALSE) +
  geom_hline(yintercept=0, linetype="dashed", color="black") +
  labs(title=myname, x="Estimated Marginal Medians and 95% HPDCI", y="AI")

```

```{r points_sf_ccdfinterval_pre, echo=FALSE, fig.width=8}

plt[["SF Trt ccdfinterval with Data Points"]] <- 
  plt[["SF Trt ccdfinterval"]] +
    geom_point(
      pch = 4,
      position = position_jitterdodge(jitter.width = 0.9, jitter.height = 0, dodge.width = 0.9, seed = 2),
      data = {
        dta_pre %>%
          group_by(RatID, Trt) %>%
          summarise(.value = mean(AI), .groups = "drop")
      },
      aes(y=.value))

print(plt[["SF Trt ccdfinterval with Data Points"]])

```

## SF AI: Contrasts:

Contrasts as a table:

```{r sf_contrasts_pre}

emmc <- rbind(contrast(emm, simple = "Trt", method = "pairwise"))
emm_show(emmc)

```

Contrasts M2 vs M3/4/5 as a median +- 95% HPDCI (HPDCI = highest posterior density continuous interval):

```{r sf_by_pre, echo=FALSE, fig.width = 8}

plt[["SF contrast"]] <- emmc %>% emm_inorder %>% 
  ggplot(aes(x = interaction(contrast, sep=", ", drop=TRUE), y = .value, fill = contrast)) %>%
  my.stat_eye() +
  geom_hline(yintercept=0, linetype="dashed", color="black") +
  labs(title=myname, x="0-2 sec", y="AI, contrast")

print(plt[["SF contrast"]])

```

# Editable PowerPoint plots generated...

```{r pptx, message=TRUE, warning=TRUE, include=FALSE, results='hide'}

myname <- "SF span=0.4 0-2 CI AI WD Contrast"

doc <- read_pptx()

for(nm in names(plt)) {
  doc <- doc %>%
    add_slide(layout = "Title and Content", master = "Office Theme") %>%
    ph_with(value = rvg::dml(ggobj = plt[[nm]]),
          location = ph_location_type(type = "body"
                                      #, width=S_width, height=S_height
                                      ),
          bg = "transparent" ) %>%
    ph_with(value = nm,
          location = ph_location_type(type = "title") )
}
doc %>% print(target = paste0(myname,".pptx"))

```
