---
title: "SF, LR dW DW | Measurement 1 (24H WD) vs Measurement 3 (SSR, Conivaptan)  groups: Bayesian regression of smoothed stretch forces"
author: "Yaromir Kobikov <kobikov.yaromir@gmail.com>"
date: "Date: `r format(Sys.time(), '%d/%m/%Y')`"
output:
  word_document: default
  html_document:
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}

source("../../BayesianPValue.R")

```

# Read SF data 

# Contra vs ipsi

The resistance was measured as the amount of mechanical work $W_{contra}$ and $W_{ipsi}$ to stretch the contra- and ipsilesional hindlimbs, where $W(T_{0-2}) = \int_{t=0}^{2} f(t)$ was stretching  force integrated over stretching distance interval from 0 to 10 mm. Asymmetry was assessed as the difference in work performed by the contra- and ipsilesional hindlimb $\Delta W = (W_{contra} - W_{ipsi})$. The control intact rat showed symmetric $W$ pattern; the $W$ for the left and right limbs were calculated instead of $W_{contra}$ and $W_{ipsi}$ for this animal.

```{r input}

my.seed <- 20231103 # reproducible Bayessian fit
my.span <- 0.4

# rds_folder_name <- "../../rds/sf/sf_0-2_CI_dW/"
rds_folder_name <- "../../rds/sf/sf_0-2_LR_dW/"

dir.create(file.path(rds_folder_name), showWarnings = FALSE)
load("../SF Sp 0-1 0.4-1 1-2 0-2 0.4-2 loess(symmetric span0.4)_rats.RData")
rm(sf) # conserve memory, we do not need the raw data

# dta <- sf_CI_dW %>%
dta <- sf_LR_dW %>%
  filter(dT %in% c("T_0_2")) %>%
  rename(Time = Period) %>% 
  droplevels(.)

```

The median (bars) and 95% HPDCI (whiskers) for SF $\Delta W$ data points. __NOT__ the uncertainties estimated by Bayesian model.

```{r plot_bars_hpdci, echo=FALSE}

myname <- "SF 0-2 span=0.4 median and 95% HPDCI all group"

# adapted from http://www.cookbook-r.com/Graphs/Plotting_means_and_error_bars_(ggplot2)/
plt[["all_group"]] <- dta %>% 
  group_by(Trt, Time) %>% 
  median_hdci(W, .width = 0.95) %>%
  ungroup() %>%
  ggplot(aes(x=interaction(Time, sep=", ", drop=TRUE), y=W, fill=Trt)) + 
    geom_bar(position=position_dodge(), stat="identity") +
    geom_errorbar(aes(ymin=.lower, ymax=.upper),
                  width=.2,   # Width of the error bars
                  position=position_dodge(.9)) +
    geom_hline(yintercept=0, linetype="dashed", color="black") +
    theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
    labs(title=myname, x="Medians and 95% HDCI", y="dW")

print(plt[["all_group"]])

```

# Analyse Measurement 1 (24H WD) vs Measurement 3 (SSR, Conivaptan) 

```{r dta_pre}
dta_pre <- dta %>% 
  filter(Trt %in% c("24H WD", "SSR", "Conivaptan")) %>% 
  droplevels(.)
```

The median (bars) and 95% HPDCI (whiskers) for SF $\Delta W$ data points. __NOT__ the uncertainties estimated by Bayesian model.

```{r plot_pre}

myname <- "SF 0-2 span=0.4 (24H WD, SSR, Conivaptan)"

plt[["Trt, Time"]] <- 
  dta_pre %>% 
  group_by(Trt, Time, dT) %>% 
  median_hdci(W, .width = 0.95) %>%
  ungroup() %>%
  ggplot(aes(x=interaction(dT, Time, sep=", ", drop=TRUE), y=W, fill=Trt)) + 
    geom_bar(position=position_dodge(), stat="identity") +
    geom_errorbar(aes(ymin=.lower, ymax=.upper),
                  width=.2,   # Width of the error bars
                  position=position_dodge(.9)) +
    geom_hline(yintercept=0, linetype="dashed", color="black") +
    theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
    labs(title=myname, x="Medians and 95% HDCI", y="dW")

print(plt[["Trt, Time"]])

```

The group sizes are (distinct rats only, same rats with different stimulation locations are counted as duplicates)

```{r input2_pre, echo=FALSE}

d.sum <- dta_pre %>% 
  arrange(Trt, Time) %>% 
  group_by(Trt, Time) %>% 
  summarise( rats = length(unique(RatID)) ) %>%
  pivot_wider(names_from=c(Trt), values_from=rats)

d.sum %>% flextable %>% fontsize(part = "header", size = 9) %>% autofit

```

## SF $\Delta W$: Bayesian regression, student family

Prior predictive check

```{r sf_check_prior_pre}

std.prior <- brm(
  data = dta_pre,
  family = gaussian,
  W ~ 0 + Trt + (1|RatID),
  prior = c(prior(normal(0, 5), class = "sigma"),
            prior(normal(0, 5), class = "sd"),
            prior(normal(0, 20), class = "b")),
  seed = my.seed, 
  chains = my.cores, 
  cores = my.cores, 
  iter = 4e4,
  sample_prior = "only",
  file = paste0(rds_folder_name, "dw_check_prior_task_2.rds"),
  file_refit = "on_change"
)


cond_eff <- conditional_effects(std.prior,
                                "Trt",
                                method="fitted")

plt[["Trt prior check"]] <- plot(cond_eff,
                                       points = TRUE,
                                       point_args = list(width = 0.1, shape="o", col="red"),
                                       plot=FALSE)[[1]] +
                                  geom_hline(yintercept=0, linetype="dashed", color="black")

set.seed(2)
plt[["Trt prior pp_check"]] <- pp_check(std.prior, ndraws = 11)

print(cond_eff[[1]])
print(plt[["Trt prior check"]])
print(plt[["Trt prior pp_check"]])

```

Model fit and QC:

```{r sf_model_summary_pre, echo=FALSE, fig.width = 10}

std.fit <-  update(std.prior, 
                   sample_prior = "yes",
                   # control = list(adapt_delta = 0.99),
                   # control = list(max_treedepth = 15),
                   seed = my.seed, 
                   cores = my.cores,
                   file = paste0(rds_folder_name, "dw_model_summary_task_2.rds"),
                   file_refit = "on_change")
summary(std.fit)
print(plot(conditional_effects(std.fit), ask=FALSE, plot=FALSE))

```

## SF $\Delta W$: MCMC conversion diagnostics

Posterior predictive check

```{r sf_MCMC_ppcheck_pre}

set.seed(2) # reproducible pp_check

plt[["SF Trt pp_check"]] <- pp_check(std.fit, ndraws = 21)
print(plt[["SF Trt pp_check"]])

```

Autocorrelations

```{r sf_MCMC_ac_pre}

plt[["SF Trt ac"]] <- stan_ac(std.fit$fit)
print(plt[["SF Trt ac"]])

```

## SF $\Delta W$: Estimated model marginals

Estimated model medians +- 95% HPDCI (HPDCI = highest posterior density continuous interval) as a table:

```{r sf_emm_pre}

emm <- emmeans(std.fit, ~ Trt, regrid = "response", nesting = NULL)
emm_show(emm)

```

Estimated model means as a median +- 95% HPDCI (HPDCI = highest posterior density continuous interval):

```{r sf_emm2_pre, echo=FALSE, fig.width = 8}

plt[["SF Trt emm"]] <- emm %>% emm_inorder %>%
  ggplot(aes(x = interaction(Trt, sep=", ", drop=TRUE), y = .value, fill = Trt)) %>%
  my.stat_eye() +
  geom_hline(yintercept=0, linetype="dashed", color="black") +
  labs(title=myname, x="Estimated Marginal Medians and 95% HPDCI", y="dW")

```

Estimated model means as a median +- 95% HPDCI ("x" - mean values of measurements for each rat)

```{r points_sf_pre, echo=FALSE, fig.width=8}

plt[["SF Trt emm with Data Points"]] <- 
  plt[["SF Trt emm"]] +
    geom_point(
      pch = 4,
      position = position_jitterdodge(jitter.width = 0.9, dodge.width = 0.9, seed = 2),
      data = {
        dta_pre %>%
          group_by(RatID, Trt) %>%
          summarise(.value = mean(W), .groups = "drop")
      },
      aes(y=.value))

print(plt[["SF Trt emm with Data Points"]])

```

The bar and whiskers plot:

```{r sf_bar_whiskers_pre, echo=FALSE}

plt[["SF Trt bar whiskers"]] <- emm %>% gather_emmeans_draws %>% 
  median_hdci(.width = 0.95) %>% 
  ungroup() %>%
  ggplot(aes(x = interaction(Trt, sep=", ", drop=TRUE), y = .value, fill = Trt)) + 
    geom_bar(position=position_dodge(), aes(y=.value),stat="identity") +
    geom_errorbar(aes(ymin=.lower, ymax=.upper),
                  width=.2,   # Width of the error bars
                  position=position_dodge(.9)) +
    geom_hline(yintercept=0, linetype="dashed", color="black") +
    theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
    labs(title=myname, x="Medians and 95% HDCI", y="dW")

print(plt[["SF Trt bar whiskers"]])

```

The ccdinterval plot with same data as above bars and whiskers:

```{r sf_ccdfinterval_pre, echo=FALSE}

plt[["SF Trt ccdfinterval"]] <- emm %>% gather_emmeans_draws %>%
  ggplot(aes(x = interaction(Trt, sep=", ", drop=TRUE), y = .value, fill = Trt)) +
  stat_ccdfinterval(point_interval = median_hdci, .width = 0.95,
             shape = 16, point_color = "black", interval_color = "black", 
             position="dodge", slab_color = NA) +
    scale_fill_brewer(palette = "Set2") + 
    scale_color_brewer(palette = "Dark2") +
    theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  # coord_cartesian(ylim = c(0, 15.5), expand=FALSE) +
  geom_hline(yintercept=0, linetype="dashed", color="black") +
  labs(title=myname, x="Estimated Marginal Medians and 95% HPDCI", y="dW")

```

The ccdinterval plot with same data as above bars and whiskers ("x" - mean values of measurements for each rat)

```{r points_sf_ccdfinterval_pre, echo=FALSE, fig.width=8}

plt[["SF Trt ccdfinterval with Data Points"]] <- 
  plt[["SF Trt ccdfinterval"]] +
    geom_point(
      pch = 4,
      position = position_jitterdodge(jitter.width = 0.9, dodge.width = 0.9, seed = 2),
      data = {
        dta_pre %>%
          group_by(RatID, Trt) %>%
          summarise(.value = mean(W), .groups = "drop")
      },
      aes(y=.value))

print(plt[["SF Trt ccdfinterval with Data Points"]])

```

## SF $\Delta W$: Contrasts:

Contrasts Trt as a table:

```{r sf_contrasts_pre}

emmc <- rbind(contrast(emm, simple = "Trt", method = "pairwise"))
emm_show(emmc)

```

Contrasts M2 vs M3/4/5 Time as a median +- 95% HPDCI (HPDCI = highest posterior density continuous interval):

```{r sf_by_pre, echo=FALSE, fig.width = 8}

plt[["SF by Trt"]] <- emmc %>% gather_emmeans_draws %>%
  ggplot(aes(x = interaction(contrast, sep=", ", drop=TRUE), y = .value, fill = contrast)) %>%
  my.stat_eye() +
  geom_hline(yintercept=0, linetype="dashed", color="black") +
  labs(title=myname, x="0-2 sec", y="dW, contrast Trt")

print(plt[["SF by Trt"]])

```


# Editable PowerPoint plots generated...

```{r pptx, message=TRUE, warning=TRUE, include=FALSE, results='hide'}

myname <- "SF span=0.4 0-2 CI dW WD vs Control"

doc <- read_pptx()

for(nm in names(plt)) {
  doc <- doc %>%
    add_slide(layout = "Title and Content", master = "Office Theme") %>%
    ph_with(value = rvg::dml(ggobj = plt[[nm]]),
          location = ph_location_type(type = "body"
                                      #, width=S_width, height=S_height
                                      ),
          bg = "transparent" ) %>%
    ph_with(value = nm,
          location = ph_location_type(type = "title") )
}
doc %>% print(target = paste0(myname,".pptx"))

```
