---
title: 'SF: Read and preprocess raw stretch force measurements 1 and 3'
author: "Yaromir Kobikov <kobikov.yaromir@gmail.com>"
date: "Date: `r format(Sys.time(), '%d/%m/%Y')`"
output:
  word_document: default
  html_document:
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}

source("../BayesianPValue.R")

```

> **Stretch‑force experiments with symmetric WD groups.**  
> Force–distance curves from both hind limbs were recorded simultaneously after bilateral peptide injection.  
> Mechanical work for each limb (WL, WR) was obtained by integrating a loess‑smoothed force curve (span = 0.4, family = "symmetric") over 0–10 mm stretch.  
> Asymmetry was quantified as the work difference ΔW = WL − WR and the log‑ratio AI<sub>L/R</sub> = log₂(WL / WR).  

## Read stretch force (SF) data 

Read raw sampled forces and match them with rat's descriptions

```{r config_vars, echo=FALSE}

PATH_DIR_ROOT <- "../data"
PATH_FOLDER_FORCE_MEASUREMENT <- "../data/Force-WD-20240513/" 
PATH_SF_DATA_CSV <- paste0(PATH_FOLDER_FORCE_MEASUREMENT, "sf_source_data_20260125.csv")

FILE_NAME = '-ForceValues.xlsx'
PATTERN = paste0('*', FILE_NAME)

DESCRIPTION_FILE_NAME <- "WD_PA 25 12 26_1_v2.xlsx"
SHEET <- "DATA_25 12 25" # for description

TEST_NUM_FILES <- 55

```

```{r}


correctDate <- Vectorize(function(date, pattern, sep="-") {
  dateSplit <- str_split(date, pattern)[[1]]
  return(paste0(dateSplit[1], sep, dateSplit[2], sep, dateSplit[3]))
})

read.StretchForce <- function(RatFile, show=TRUE) {
  ratName <- sub(FILE_NAME, "", basename(RatFile), fixed = TRUE)
  ratNum <- str_split(ratName, "SDU")[[1]][2]
  ratName <- paste0("SDU", "-", ratNum)
  sheetNames <- RatFile %>% 
    excel_sheets() %>% 
    as_tibble %>% 
    filter(grepl(pattern = "F([0-9]+)", value))
  
  df <- 
    map_df(
      .x = sheetNames$value,
      ~read_excel(path = RatFile, sheet = .x, range = cell_cols("A:C")),
      .id = "replication"
      ) %>% 
    mutate(RatID = ratName) %>% 
    relocate(RatID, .before = replication)
  
  if(show) print(paste0(ratName, " -> ", toString(count(sheetNames)$n), ", ", count(df)$n))
  return(df)
}

my.rwData <- function() {
  
  # if(file.exists(PATH_SF_DATA_CSV)) {
  #   print(paste0("File ", " - ", PATH_SF_DATA_CSV, " - exists"))
  #   return(read_csv(PATH_SF_DATA_CSV, show_col_types = FALSE))
  # }
  print("In process...")
  
  sf <- 
    map_df(
      .x = list.files(path = PATH_FOLDER_FORCE_MEASUREMENT, pattern = PATTERN, full.names = TRUE),
      ~read.StretchForce(.x)
    )
  
  if(length(unique(sf$RatID)) == TEST_NUM_FILES) print(paste0(TEST_NUM_FILES, " - ", "OK"))
  
  # match rat's description with raw force measurements
  rats_description <-
    file.path(PATH_DIR_ROOT, DESCRIPTION_FILE_NAME) %>% 
    read_excel(SHEET, .name_repair = "universal", na = c("", "NA", "NULL")) %>%
    rename(
      BW_pre        = BW..pre.,
      BW_post       = BW..post.,
      Measurement.1 = Measurement.1,
      Measurement.3 = Measurement.3
    ) %>%
    select(
      RatID,
      Measurement.1,
      Measurement.3,
      BW_pre, BW_post,
      MP.1.1, MP.1.2, MP.1.3,
      MP.3.1, MP.3.2, MP.3.3
    ) %>%
    rename_with(
      ~ sub("^MP\\.(\\d+)\\.(\\d+)$", "PA\\1.\\2", .x),
      .cols = matches("^MP\\.(1|3)\\.(1|2|3)$")
    ) %>% 
    mutate(RatID = factor(RatID),
           Trt1 = factor(Measurement.1, c("24H WD", "Control 1")),
           Trt3 = factor(Measurement.3, c("Sp", "Control 1", "Control 2", "Control 3", "SSR", "Conivaptan"))
    )
  
  sf <-
    merge(sf, rats_description, by=c("RatID"), all.x=TRUE, all.y=TRUE, sort=FALSE) %>%
    filter(replication %in% c(1:3, 13:15)) %>%
    pivot_longer(
      cols = c("Left", "Right"),
      names_to = c("Side"),
      values_to = "SF"
    ) %>%
    filter()
    mutate(
      SF = SF*(-1),
      rep_i = as.integer(replication),
      Period = case_when(
        rep_i %in% 1:3   ~ "24H",
        rep_i %in% 13:15 ~ "2_3H",
        TRUE ~ NA_character_
      ),
      Period = factor(Period, levels = c("24H", "2_3H")),
      Trt = case_when(
        rep_i %in% 1:3   ~ paste0("24H.",  Trt1),
        rep_i %in% 13:15 ~ paste0("2_3H.", Trt3),
        TRUE ~ NA_character_
      )
    )

  write_csv(x = sf, PATH_SF_DATA_CSV)
  return(sf)
}

```



```{r input, results='hide'}

myspan <- 0.4
myname0 <- paste0("SF loess(symmetric span=", myspan, ")")

data <- my.rwData()

sf <-
  data %>%
  rename(repN = replication) %>%
  mutate(
    RatID = factor(RatID),
    # MS = factor(MS, c("Left", "Right"), c("L", "R")),
    # Later = factor(OperationSide == MS, c(FALSE, TRUE), c("Contra", "Ipsi")),
    # Trt = paste0(Trt, ".", OperationSide),
    Trt = factor(Trt),
    repN = factor(as.integer(repN))
  )

# sf[!complete.cases(sf), c("RatID", "Op3", "Trt", "repN", "Later", "SF")] # Any unmatched rats? Any unmeasured forces?
```

The group sizes are (distinct rats only, same rats with different stimulation locations are counted as duplicates)

```{r dataset_n, echo=FALSE}

d.sum <- sf %>% 
  arrange(Trt, Period) %>% 
  group_by(Trt, Period) %>% 
  summarise(rats = length(unique(RatID))) %>%
  pivot_wider(names_from=c(Trt), values_from = rats)

d.sum %>% 
  flextable %>% 
  fontsize(part = "header", size = 9) %>% 
  autofit

```

## SF, 0-2 sec: Visual inspection of smoothing

<!-- Filter 0-2 sec measurements, taking extra 0.5s for better smoothing. -->

```{r SF_combo_dta}

mydta <- sf %>%
  filter(!is.na(SF)) %>%
  droplevels(.) %>%
  mutate(figName = paste0(RatID, ", ", Trt, ", replicate ", repN)) %>%
  select(figName, Period, Time, Side, Trt, SF) %>%
  # filter(figName == "SDU-115, 3h.BNI.L, replicate 12") %>%
  group_by(figName)

```

Stretch force figure combining Contra, Ipsi and their difference.

```{r example_12_rep, echo=FALSE, fig.width=12, fig.height=8}

myspan <- 0.4

name_fig <- "SDU-005, 24H.24H WD, replicate 1"
name_plt <- paste0("raw ", name_fig)

plt[[name_plt]] <- 
  mydta %>%
  filter( figName == name_fig) %>%
  # pivot_wider(names_from = Later, values_from = SF) %>%
  # mutate( 
  #   "Contra - Ipsi" = Contra - Ipsi
  # ) %>%
  # pivot_longer(c("Contra", "Ipsi", "Contra - Ipsi"), names_to = "Limbs", values_to = "SF") %>% 
  ggplot(aes(x = Time, y = SF, group = Side, col = Side)) + 
  geom_line() + 
  geom_smooth(
    formula = y ~ x, 
    aes(col = paste0("loess ", myspan)), 
    method ="loess", 
    span = myspan, 
    se = FALSE, 
    method.args = list(family="symmetric")
  ) +
  labs(
    title = paste0(myname0, ", ", name_fig), 
    x = "Time (sec)", 
    y = "Stretching force (mg)"
  )

print(plt[[name_plt]])

```

```{r}

doc <- read_pptx()

for(nm in names(plt)) {
  doc <- doc %>%
    add_slide(
      layout = "Title and Content",
      master = "Office Theme"
    ) %>%
    ph_with(
      value = rvg::dml(ggobj = plt[[nm]]),
      location = ph_location_type(
        type = "body"
        #, width=S_width,
        # height=S_height
        ),
      bg = "transparent"
    ) %>%
    ph_with(
      value = nm,
      location = ph_location_type(type = "title")
    )
}

doc %>% print(target = paste0("SF loess(symmetric span0.4)", ".pptx"))

```

Visually check the smoothing by loess(), varying the "span" to see if alternatives are better than our choice. 

```{r check_loess, include=FALSE}

myspan.test <- 0.4

junk <- mydta %>%
  group_map( ~{
    theName <- paste(.y)
    ggplot(data=.x, aes(y = SF, x = Time, group = Side, col = Side)) +
      geom_line() +
      geom_smooth(formula = y ~ x, aes(col = paste0(myspan.test + 0.1)), method="loess", span = myspan.test + 0.1, se=FALSE, method.args=list(family="symmetric")) +
      geom_smooth(formula = y ~ x, aes(col = paste0(myspan.test - 0.1)), method="loess", span = myspan.test - 0.1, se=FALSE, method.args=list(family="symmetric")) +
      geom_smooth(formula = y ~ x, aes(col = paste0(myspan.test)), method="loess", span=myspan.test, se=FALSE, method.args=list(family="symmetric")) +
      # theme(legend.position = "none") +
      labs(title=paste(myname0, theName), x="Time (sec)", y="Stretching force (mg)")
  } ) %>%
  setNames(unique(sort(mydta$figName)))

mydta %>% n_groups()

```

```{r write_pptx_check_loess, echo=FALSE, fig.width=12, fig.height=8}

for(nm in unique(mydta$figName)) {
  print(junk[[nm]])
}

doc <- read_pptx()

for(nm in unique(mydta$figName)) {
  doc <- doc %>%
    add_slide(
      layout = "Title and Content",
      master = "Office Theme"
    ) %>%
    ph_with(
      value = rvg::dml(ggobj = junk[[nm]]),
      location = ph_location_type(
        type = "body"
        #, width=S_width,
        # height=S_height
        ),
      bg = "transparent"
    ) %>%
    ph_with(
      value = nm,
      location = ph_location_type(type = "title")
    )
}
doc %>% print(target = paste0(myname0, "_0.4_.pptx"))

```

Define smoother for all Time points.

```{r smoother_W_AI}

# 0-1 0.4-1 1-2 0-2 0.4-2

smoother_W <- function(dta, ...) {

  # dtaF <- dta %>% filter(Time < 10.5, !is.na(SF))

  dtaF <- dta %>% filter(!is.na(SF))

  if(nrow(dtaF) < 1 | ncol(dtaF) < 2) {
    print(dta)
    return(data.frame())
  }

  my.sm <- loess(SF~Time, data=dtaF, family="symmetric", span=myspan)

  i_0_1  <- integrate(function(x) predict(my.sm, newdata=x), 0, 1)
  i_0_4_1  <- integrate(function(x) predict(my.sm, newdata=x), 0.4, 1)
  i_1_2 <- integrate(function(x) predict(my.sm, newdata=x), 1, 2)
  i_0_2 <- integrate(function(x) predict(my.sm, newdata=x), 0, 2)
  i_0_4_2 <- integrate(function(x) predict(my.sm, newdata=x), 0.4, 2)

  data.frame(
    dT = c(
      "T_0_1",
      "T_0.4_1",
      "T_1_2",
      "T_0_2",
      "T_0.4_2"
      ),
    W = c(
      i_0_1$value,
      i_0_4_1$value,
      i_1_2$value,
      i_0_2$value,
      i_0_4_2$value
      )
  )
}

```

Remove unneded columns from SF dataset

```{r clean_SF}

sf_ai_dw <-
  sf %>%
  # filter(RatID == "TBI-250", Trt %in% c("UBI.R")) %>%
  select(Trt, Period, RatID, repN, Side, Time, SF) %>%
  ungroup %>%
  arrange(Trt, RatID, repN, Side, Time)

```

Compute work (W) as integrals (0-1 sec, 0.4-1 sec, ...) of smoothed force measurements. dW = W_Left - W_Right

```{r compute_W_LR}

sf_LR_dW <- sf_ai_dw %>%
  group_by(Trt, Period, RatID, repN) %>%
  pivot_wider(names_from = Side, values_from = SF) %>%
  # spread(MS, SF) %>%
  mutate( SF = Left - Right ) %>%
  group_modify(smoother_W) %>%
  ungroup %>%
  droplevels(.)

```

Compute AI = log2(Left/Right) using integrals (0-1 sec, 0.4-1 sec, ...) of smoothed force measurements.

```{r compute_AI_LR}

sf_LR_AI <- sf_ai_dw %>%
  group_by(Trt, Period, RatID, repN, Side) %>%
  group_modify(smoother_W) %>%
  ungroup %>%
  group_by(RatID, Trt, repN, dT) %>%
  pivot_wider(names_from = Side, values_from = W) %>%
  # spread(MS, W) %>%
  mutate( AI = log2( Left / Right ) ) %>%
  ungroup %>%
  droplevels(.)

```

Save preprocessed SF

```{r smooth}

myname <- paste0("SF Sp 0-1 0.4-1 1-2 0-2 0.4-2 loess(symmetric span", myspan, ")")

save(sf_LR_dW, sf_LR_AI, sf, file=paste0(myname, "_rats.RData"))

```
